Exception Handling Bug Hazards in Android

### 2016-04-20 SPRINGER_JOURNAL/msr_journal.tex
addressing ref 3 and Georgios comments

\emph{\textbf{Identifying the Concerns Related to Root Exceptions.}} To get a broader view of the root exceptions of stack traces,
we performed a manual inspection in order to identify the underlying
concerns related to the most frequently reported root exceptions.
Besides the exceptions related to programming mistakes mentioned before, we also looked for exceptions related to concerns that are known as sources of faults in mobile development: concurrency~\cite{ama2012} backward compatibility~\cite{McDon13}, security~\cite{enck2011study,was2010} and resource management (IO, Memory, Battery)~\cite{Zhang12}. Since it is infeasible to inspect the code responsible for throwing every exception reported in this study,
the concern identification for each exception was based on intended
meaning of the particular exception type, as defined in
its Javadoc documentation and in the Java specification.
For example: (i) an instance of ArrayOutOfBoundException
refers to a programming mistake according to its Javadoc; and (ii) the Java specification lists all
exceptions related to backward compatibility~\footnote{http://docs. oracle.com/javase/specs/jls/se7/html/jls-13.html}, such as
InstantiationError, VerifyError, and IllegalAccessError.

@@

\emph{\textbf{Checked Exception wrapping an Error.}} Most of these wrappings
 were also caused by the reflection library used by applications' methods. The methods responsible for the wrappings
were also native methods written in C. Table~\ref{tab:exampeswrap} illustrates
some of these wrappings --- some of them are masking an OutOfMemoryError
into a checked exception.
\textcolor{red}{By confronting the API user with a checked exception,
the API designer is forcing the client to handle the exceptional condition.
In this case, the exception condition masks an unrecoverable error,
which makes the task of providing a proper handler
infeasible.\footnote{TODO: Please note that Errors are not supposed to be
caught. The main problem here is that the program continues with e.g. no
memory available.}
This kind of wrapping is even more dangerous the previous one
and may lead to the ``exception confusion'' described next.}

@@

To make the analysis easier, we further refine this question into sub-questions, such as: \emph{How often do developers handle exceptions?} \emph{How often do developers throw exceptions?} And, to show us how developers may deal with the exception handling (EH) code: \emph{Do developers know about Java EH best practices and/or Android specific EH best practices?}

@@

%\emph{\textbf{RQ 2.5:}  Are developers aware of the non-declared checked exceptions?}

%\emph{RQ 2:Have developers ever faced a crash caused by NullPointerException?}
%\emph{RQ 3:Are developers aware of the robustness threats caused by cross-type wrappings?}; %\emph{RQ 4: Are developers aware of the robustness threats caused by JNI undocumented checked exceptions?}

@@

\textcolor{red}{Since our aim is to learn from a large number of developers, we
use surveys which are known to scale well.}
\textcolor{red}{The survey is split into two logical sections; the motivations behind each logical section are:
(i) questions aiming at learning from developers about the usage of the exception handling code in app development;  and
(ii) questions focusing on getting developers perceptions about the bug hazards detected in the first phase of this study.
The survey comprises multiple choice or Likert-scale questions and open-ended questions. The multiple choice questions are intermixed with open-ended questions to further elicit the developer's opinions. Moreover, the survey also contains Likert-scale questions to force participants to make a choice. Overall, the survey includes 13 open-ended questions, 5 Likert-scale questions and 10 multiple choice questions. The respondents could complete the survey in about 15 minutes.}

\textcolor{red}{Given~\footnote{TODO: I (GG) would put this paragraph before
the previous one} the exploratory nature of our research, we used methods from Grounded Theory ~\cite{charmaz2006} to answer some of our research questions by iterating through the open-ended survey responses. The grounded theory coding used in this study consists of at two phases: (1) initial coding entails a close reading of the data and (2) later we used focused coding ~\cite{charmaz2006} to pinpoint and develop the most salient themes ~\cite{charmaz2006} in the analyzed data. Answers to different questions in the survey were coded separately. For all open-ended questions in the survey, coding was done by two researchers until saturation was reached. Disagreements were discussed and resolved as part of the coding, thus, we are unable to report level of agreement. In all cases, the first author was one of the two coders, and one of the other authors coded the data as well in close collaboration with the first author until saturation was reached. The first author then went through the remaining responses to assign codes to responses.}

@@

\textcolor{red}{The process to
identify the type of exceptions reported in stack traces may not be completely
accurate.
Specifically, we performed the selection of the version of the exception source code or bytecode
to analyze, as follows: (i) for all exceptions signaled by the Android Platform
and the Java Environment the analysis was based on Version 4.4 of Android platform (API level 19);  (ii)
for the exceptions signaled by applications, the analysis considered
the last version of the application source-code available on the Github/Google code; and (iii)
for all exceptions signaled by third-party libraries the type analysis
considered the latest bytecode version available on the app repository.
When the exception could not be found automatically or manually (based on the chosen version),
we classified the exception as ``Undefined''.  Only 31 exceptions
remained undefined, which occurred in 60 different exception stack traces out of 6,005 mined stack traces.
Hence, the exception type could not be accurately identified in 1\% of the mined
exception flows.
There should be situations where the exception type (i.e., checked or runtime) changes from version to another.
Although we believe that it will not happen very often, we did not investigate
this issue.}

@@

\begin{acknowledgements}
This work is partially supported by INES, funded by CNPq and
FACEPE, grants 573964/2008-4 and APQ-1037-1.03/08.
\end{acknowledgements}

>>>>

\emph{\textbf{Identifying the Concerns Related to Root Exceptions.}} To get a broader view of the root exceptions of stack traces,
we performed a manual inspection in order to identify the underlying
concerns related to the most frequently reported root exceptions.
Besides the exceptions related to programming mistakes mentioned before, we also looked for exceptions related to concerns that are known as sources of faults in mobile development: concurrency~\cite{ama2012} backward compatibility~\cite{McDon13}, security~\cite{enck2011study,was2010} and resource management (IO, Memory, Battery)~\cite{Zhang12}. Since it is infeasible to inspect the code responsible for throwing every exception reported in this study,
the concern identification for each exception was based on intended
meaning of the particular exception type, as defined in
its Javadoc documentation and in the Java specification.
For example: (i) an instance of ArrayOutOfBoundException
refers to a programming mistake according to its Javadoc; and (ii) the Java specification lists all
exceptions related to backward compatibility
~\footnote{http://docs.oracle.com/javase/specs/jls/se7/html/jls-13.html}, such as
InstantiationError, VerifyError, and IllegalAccessError.

@@

\emph{\textbf{Checked Exception wrapping an Error.}} Most of these wrappings
 were also caused by the reflection library used by applications' methods. The methods responsible for the wrappings
were also native methods written in C. Table~\ref{tab:exampeswrap} illustrates
some of these wrappings --- some of them are masking an OutOfMemoryError
into a checked exception.
\textcolor{red}{On the one hand, by confronting the API user with a checked exception,
the API designer is forcing the client to handle the exceptional condition.
On the other hand, according to Java specification Errors are not supposed to be
caught. In this case, even if the exception is caught by a handler, the problem 
that trigged the Error remains: there is not enought memory to execute the app.
This kind of wrapping is even more dangerous than the previous one
and may lead to the ``exception confusion'' described next.}

@@

To make the analysis easier, we further refine this question into sub-questions, as follows. To investigate whether the development of exception handling code is a daily concern for developers or it is something that developers face seldom and for that reason do not care much, the following subquestions are added: \emph{How often do developers handle exceptions?} \emph{How often do developers throw exceptions?} And, to check if the Java EH best practices, available in literature, are used in practice and to investigate if any Android-specific EH best practice are used by developers, the following subquestion is added  subquestion: \emph{Do developers know about Java EH best practices and/or Android specific EH best practices?}

@@

@@

%~\footnote{TODO: I (GG) would put this paragraph before the previous one} 
%~\footnote{TODO: I (RSC) merged the beginning of both paragraphs. }

\textcolor{red}{Given the exploratory nature of our research, we used methods from Grounded Theory ~\cite{charmaz2006} to answer some of our research questions. Since our aim is to learn from a large number of developers, we use surveys which are known to scale well. The survey is split into two logical sections; the motivations behind each logical section are: (i) questions aiming at learning from developers about the usage of the exception handling code in app development;  and (ii) questions focusing on getting developers perceptions about the bug hazards detected in the first phase of this study.}

\textcolor{red}{The survey comprises multiple choice or Likert-scale questions and open-ended questions. The multiple choice questions are intermixed with open-ended questions to further elicit the developer's opinions. Moreover, the survey also contains Likert-scale questions to force participants to make a choice. Overall, the survey includes 13 open-ended questions, 5 Likert-scale questions and 10 multiple choice questions. The respondents could complete the survey in about 15 minutes.}

\textcolor{red}{We used grounded theory coding to iterate through the open-ended survey responses. The grounded theory coding used in this study consists of at two phases: (1) initial coding entails a close reading of the data and (2) later we used focused coding ~\cite{charmaz2006} to pinpoint and develop the most salient themes ~\cite{charmaz2006} in the analyzed data. Answers to different questions in the survey were coded separately. For all open-ended questions in the survey, coding was done by two researchers until saturation was reached. Disagreements were discussed and resolved as part of the coding, thus, we are unable to report level of agreement. In all cases, the first author was one of the two coders, and one of the other authors coded the data as well in close collaboration with the first author until saturation was reached. The first author then went through the remaining responses to assign codes to responses.}

@@

\textcolor{red}{The process for
identifing the type of exceptions reported in stack traces may not be completely
accurate.
Specifically, we performed the selection of the version of the exception source code or bytecode
to analyze, as follows: (i) for all exceptions signaled by the Android Platform
and the Java Environment the analysis was based on Version 4.4 of Android platform (API level 19);  (ii)
for the exceptions signaled by applications, the analysis considered
the last version of the application source-code available on the Github/Google code; and (iii)
for all exceptions signaled by third-party libraries the type analysis
considered the latest bytecode version available on the app repository.
When the exception could not be found automatically or manually (based on the chosen version),
we classified the exception as ``Undefined''.  Only 31 exceptions
remained undefined, which occurred in 60 different exception stack traces out of 6,005 mined stack traces.
Hence, the exception type could not be accurately identified in 1\% of the mined
exception flows.
There should be situations where the exception type (i.e., checked or runtime) changes from version to another.
Although we believe that it will not happen very often, we did not investigate
this issue.}

@@

\begin{acknowledgements}
This work is partially supported by the National Institute of Science and Technology for Software Engineering (INES), CNPq and FACEPE, grants 573964/2008-4, 552645/2011-7, and APQ-1037-1.03/08, CNPq Universal grant 484209/2013-2, and CAPES/PROAP.
\end{acknowledgements}
